\section{Sound Speed Profile Estimation}

Although the speed of sound $c$ in water is considered constant at around $1500 \si{\meter}/\si{\second}$, it can vary when considering environments of both small and large scales, respectively in small and large proportions \cite{kieffer_sound_1977,trusler_determination_2017,li_sound_2015,pinson_sound_2010}. Many properties can affect the speed of sound in water, such as pressure, temperature, salinity, and water currents \cite{wilson_speed_1959,medwin_fundamentals_1998,mcdougall_getting_2011,delgrosso_new_1974}. As these properties vary over the body of water, so does the velocity, creating a sound speed profile (SSP) in the body of water.

Initial works in the labor of approximating the velocity profile in water aimed at estimating these key properties, as well as the relationship between them and the velocity of water. Works such as \cite{wilson_speed_1959,medwin_fundamentals_1998,mcdougall_getting_2011,delgrosso_new_1974,chen_speed_1977,wong_speed_1995,roquet_accurate_2015} worked under this framework, with some still being used today to model the behavior of water currents on a global scale.

In \cite{wilson_speed_1959}, their experiments showed that (in distilled water) $c$ increases with pressure, and also with temperature up to a certain point, either flattening or decreasing with higher temperatures. In \cite{medwin_fundamentals_1998}, empirical models for the $c$ are provided, also taking salinity into account in their formulation.

The TEOS-10 model \cite{mcdougall_getting_2011} is another method for estimating the velocity profile given the temperature, pressure, and salinity parameters in the body of water, solving chemical-thermodynamics equations to find the adequate profile.

The pursuit of knowing the speed of sound at each point of a body of water -- also known as sound speed profile, or SSP -- allows the knowledge of the properties of a wave that travels through this medium. By Snell's law, a change in speed of a wave also changes its direction, therefore a wave that travels through a sufficiently large body of water will be also transformed within it, resulting in curves, refractions, reflections, and formation/dissolution of beams.

By knowing the SSP, we can know the path a unidirectional wave would course. By treating an omnidirectional wave as a cluster of unidirectional waves, we can calculate the travel-time between the source and any point within the medium, as well as its \qt{shell}s profile.

Most modern approaches avoid estimating these properties, and treat this SSP estimation as an inversion problem. That is, since the knowledge of the SSP allows calculating the path of a wave through the medium, then the knowledge of these paths permits the estimation of the SSP in the medium. Different techniques tackle this inversion problem by using measurements, and employing the acquired information through different lenses.

Most techniques utilize sound sources and receptors within the body of water, and by comparing the properties of the waves that are produced and received, estimate the path of the waves; and from this, the sound speed profile of the water body.

\subsection{Methods for SSP inversion}

Many different methods can be used for estimating the sound speed profile, using different information measured in sensors along the body of water. These data can be the travel-time between source and receptors, the measured pressure in the receivers, difference in phase and amplitude between the devices, or even the temperature-salinity-pressure used in the previously cited methods. Along with this, the techniques can use these measurements through either direct SSP estimation methods, as well as via machine learning.

\subsubsection{Travel-time tomography}

Travel-time tomography (TTT) \cite{bormann_seismic_2012,stefanov_travel_2019} is one of the simplest methods for SSP estimation. It uses the difference in arrival time of the acoustics signals in the sensors, requiring a spatial range of sources and receiver(s).

In general, it discretizes the SSP in layers, where each layer behaves having a constant velocity. Then, the wavepath's time in each layer is calculated, resulting in a system of equations that is used to estimate the velocity in each layer. It can also work with assumed functions for the SSP, ditching the layering assumption, but increasing the problem's complexity.

Its simpleness, robustness and computational efficiency are its main advantages, the later allowing real-time implementation. However, it depends on a precise positioning and syncing of the devices, and its discretization assumption results in limited resolution.

\subsubsection{Matched field processing and inversion}
In matched field processing (MFP) \cite{gemba_adaptive_2017} and matched field inversion (MFI) \cite{dosso_bayesian_2011}, the measured pressure data in the receivers is compared to acoustic pressure models for different SSPs, matching and adapting the models to the data through optimization algorithms in order to find the most similar one.

These techniques are similar to beamforming, which are used in many fields in acoustics and signal processing \cite{schmidt_environmentally_1990,brienzo_broadband_1993}, as they use spatial correlation between the receivers and their measurements, as well as model the acoustic field through them.

MFP and MFI also employ modal decomposition of the signals, with each mode having different behavior in water. They incorporate more complex information regarding the signals, such as amplitude and phase, depending not only in travel time. Among their advantages, we have a better use of measured information, enhanced spatial resolution -- compared to TTT -- and improved robustness to noise due to coherent processing. However, it requires spatial-temporal information on the source and receivers, such as directivity and spectral behavior. It also isn't very suitable for very large bodies, given that a larger body would result in an exponentially more complex acoustic pressure model; and is more computationally complex, as well as susceptible to local minima.

\subsubsection{Full waveform inversion}

Full waveform inversion (FWI) techniques \cite{virieux_overview_2009,fichtner_multiscale_2013} aim the waveform reconstruction of the signals present in the SSP, considering many different behaviors that are neglected by the aforementioned techniques. Here, the error between the estimated and measured waveforms is minimized, being implemented in either time- or frequency-domains.

Given this, the FWI technique is among the ones that maximally use the receivers' available information, and resulting in the most faithful SSP, as well as maximally using the available signals. However, due to considering complex signal interactions and multiple wavepaths, there is a significant increase in the mathematical and computational complexity. This, coupled with the non-linearity of the problem, results in a very intense and sometimes unfeasible technique, that also requires a decent initial parameter estimation for its robustness and convergence.

\section{Fundamentals of travel-time tomography}

As previously explained, Travel-Time Tomography (TTT) uses the delay -- travel time -- between the wave's generation at the source and its arrival in the receptors, to estimate the path traveled, and therefore the weave's speed along the route taken.

\input{figures/1.1-WavepathHeterogeneous.tex}

Consider a wave that follows the path from \cref{fig:sec2:wavepath_heterogeneous-variable_environment}, this being represented by the curve $\bvr[v]$. Assuming that the speed of sound $c(\bvp)$ is a function of space ($\bvp = [x,y,z]$), then the travel time for the curve $\bvr[v]$ is
\begin{equation}
	t_{\bvr[v]} = \int_{\bvr[v]} \frac{1}{c(\bvp)} \dd\bvp.
\end{equation}

The objective is to, given a set of $N$ measures $t_{\bvr[v]{i}}$, these representing different wave paths (from different sources, or being read in different sensors), estimate the sound speed profile (SSP) $v(\bvp)$ at a discrete or continuous set of points.

Many different approaches can be taken to solve this problem, including to discretize space, approximate the SSP as polynomial through Spline or least-squares fitting, among others. This technique can also be employed in 1D, 2D or 3D situations, and getting spatial diversity through either (or both) multi-source or multi-receiver means.

\subsection{Papers and articles' review}

As stated previously, this technique is versatile due to its simplicity, being employable in a range of applications from acoustics seismology to computarized living-tissue tomography and electromagnetic evaluation of water properties.

\subsubsection{Aki, K - 1976, 1977}

The works by Aki in \cite{aki_determination_1976,aki_determination_1977} are sometimes regarded as the pioneering papers in travel-time tomography. Their work is on the field of seismology and geological layer classification through the SSP estimation.

\textbf{Objective} - The purpose of this paper is to use travel-time information on different geologic layers to estimate the SSP on them, as well as guiding earthquake origin/epicenter estimation.

\textbf{Data and assumptions} - The data used is the travel-time from sub-surface waves generated by earthquakes. For such, the media (geologic layers) are discretized among them, as well as each layer being separated into cells, with varying velocities within each cell. The environment is assumed at least 2D, as the media is separated into layers and then into cells.

\textbf{Techniques used} - Each layer is assumed to have an average velocity, and each cell in that layer deviates from this average. Least-squares technique is used: the model is adjusted to match the modeled and measured first arrival times.

\subsubsection{Ali - 2019}

This paper \cite{ali_opensource_2019} aims to estimate the SSP on different tissues via ultrasound computer tomography. They use a joint Eikonal equation and ray tracing solutions, coupled with different regularization techniques for smoothness constraints.

\textbf{Objective} - Use travel-time information to reconstruct the SSP, allowing for a better imaging process for biological tissue.

\textbf{Data and assumptions} - Travel-time information from ultrasound CTs, with a ring of receivers around the object.

\textbf{Technique used} - The tissue is considered to be split into 2D cells. An iterative minimization process is employed, searching concomitantly for the SSP of the tissue, as well as the path taken by the emitted signals through the tissue. The Fast Marching Method (FMM) \cite{chopp_improvements_2001,sethian_3d_1999,sethian_fast_1999} is used for the raytracing portion, and an iterative least-squares minimization to find most appropriate speed on each cell in the tissue.

\textbf{Similar paper} - The paper by Hormati \cite{hormati_robust_2010} uses a similar approach to that in \cite{ali_opensource_2019}. Their objective is to determine the velocity profile in an ultrasound tomography by discretizing the tissue; applying a (discrete version of) the travel-time integral to the system; and estimating the system parameters, as well as the ray-lengths within each discrete cell, to iteratively achieve the tissue's SSP. They apply a few different techniques for the iterative minimization process, but this contrast is outside the TTT-SSP scope.

\textbf{Similar paper} - The paper by Jeong \cite{jeong_investigating_2024} uses a similar approach to that in \cite{hormati_robust_2010}, comparing TTT to reflection tomography, and employing them in convolutional neural network (CNN) deep learning training.

\textbf{Similar paper} - In this paper by Hellmann \cite{hellmann_borehole_2023}, they use travel-time tomography for estimating the SSP in glacier boreholes, along with a model correction step for the positions of the sources/receivers along the boreholes.

\subsubsection{Philips - 1991}

This paper \cite{phillips_traveltime_1991} studies different methods for discrete travel-time tomography, comparing their results in metrics such as roughness, misfit, and spread. These methods trade overall performance against that of some metric.

\textbf{Objective} - Compare different inverse paradigms in achieving the parameters for the SSP model.

\textbf{Data and assumptions} - Synthetic discretized data. There are 100 sensors/receivers in total, 25 per side of the square-grid of sensors. Each device sends array to all other devices. Ray-bending is disregarded.

\textbf{Technique used} - Similar to the previously studied papers, the 2D environment is split into cells, and inversion of the slowness model is used to achieve the results. Different conditions on the inversion are considered, such as smoothing, regularization, dampening, and iterative reweighing. These different techniques are compared for the proposed metrics, exploring their advantages for different applications.

\subsubsection{Zhang - 2017}

This paper \cite{zhang_nonlinear_1998} proposes an adaptation of the travel-time tomography approach to include reflections as well. Their modeling minimize a combination of average slowness (travel-time $/$ path length) and apparent slowness (derivative of travel-time w.r.t. surface (horizontal) distance), along with a regularization term.

\textbf{Objective} - Achieve a TTT model that incorporates reflections alongside direct-path waves, in estimating the parameters of the system.

\textbf{Data and assumptions} - It assumes that reflected waves can arrive in the sensors at similar times as the direct-path waves, enabling their coupled utilization in estimating the system.

\textbf{Technique used} - A conjugate gradient minimization approach is employed, as is usual in TTT inversion problems. Also, similarly to previously explored papers, a cell-based approach is taken.

\subsubsection{Dines - 1979}

This paper \cite{dines_computerized_1979} also works in TTT, now estimating both the velocity and attenuation fields from the measured data. This paper works with electromagnetic signals, however the mathematics and theories are the same.

\textbf{Data and assumptions} - They assume an exponential decay on the signals' amplitudes over time, as well as straight waves, dismissing refraction effects. The same media discretization is done, separating it into 2D cells along the depth and distance between the columns of sources/receivers.

\textbf{Technique used} - An iterative minimization is done over all sources and sensors (and their pair-wise paths), searching for either the velocity or attenuation fields. For solving the ill-posed problem, they propose an approach that considers one ray at a time, updating all velocities in the cells that are traversed by the current ray, until sufficient convergence is achieved.

\subsubsection{Tang - 2024}

This paper \cite{tang_travel_2024} proposes the use of the Fast Sweeping Method (FSM) instead of FMM for solving the Eikonal equations during the TTT inversion process. They present the FSM method and derive its implementation for TTT, as well as an extensive comparison between the two solvers for the Eikonal problem. Modeling of reflections is also discussed in the paper.

%\textbf{Objective} - Implement a FSM-based TTT technique for SSP estimation.

\textbf{Data and assumptions} - Synthetic environment and signals are considered, for a wide gamma of situations.

\textbf{Technique used} - The paper's focus is on exploring the use of FSM instead of FMM, and not so much in the TTT process, or other details surrounding it.

\subsubsection{Common themes between articles}
\NewDocumentCommand{\citep}{m}{%
	\citeauthor*{#1}, \citeyear{#1}%
}
\def\X{\ensuremath{\mathsf{X}}}

The table below compiles a comprehensive list of the themes that are commonly approached among the papers presented previously, as well as a short commentary on the different features or considerations, when relevant.
\begin{table}[H]\small\centering
	\begin{tabular}{l c c c c c l}
		Paper 								& Discrete 	& Ray bending 	& CG 	& $P_{\alpha}$ 	& $\bvB{i}$ & Extra 	\\
		\citep{aki_determination_1976} 		& \X 		& -- 			& --	& \X			& --		& --		\\
		\citep{aki_determination_1977} 		& \X 		& -- 			& --	& \X			& --		& Perturbation analysis 		\\
		\citep{ali_opensource_2019}			& \X		& \X			& \X	& \X			& \X		& Presented a statistical linearization \\
		\citep{hormati_robust_2010}			& \X 		& \X			& \X	& \X			& --		& --		\\
		\citep{jeong_investigating_2024}	& \X		& \X			& --	& \X			& --		& --		\\
		\citep{hellmann_borehole_2023}		& \X		& \X			& --	& \X			& --		& Position estimation	\\
		\citep{phillips_traveltime_1991}	& \X		& --			& --	& \X			& \X		& Compares many techniques	\\
		\citep{zhang_nonlinear_1998}		& \X		& \X			& \X	& \X			& --		& Considers reflections	\\
		\citep{dines_computerized_1979}		& \X		& --			& --	& --			& --		& Considers attenuation; Works ray-by-ray \\
		\citep{tang_travel_2024}			& \X		& \X			& \X	& --			& --		& Uses FSM instead of FMM
	\end{tabular}
\end{table}

\subsection{Standard discrete travel-time tomography approach}

All of the cited papers took an approach similar to the one established below. Differences may arise from ignoring ray bending, type of regularization in the minimization process, amount and positioning of sources and receivers, final purpose for the TTT, or some other particularization or assumption.

In these regards, the most complete papers are the ones by Ali \cite{ali_opensource_2019}, Phillips \cite{phillips_traveltime_1991}, and Hormati \cite{hormati_robust_2010}, where they consider regularization, ray bending, and a large amount of sources and receivers. We now will lay the common ground among all the presented papers, the median modeling considerations and formulations, as well as some important distinctions and particularities that are shared between some of them. Note that the framework presented here isn't identical to the ones in the cited papers \cite{aki_determination_1977,aki_determination_1976,ali_opensource_2019,hormati_robust_2010,jeong_investigating_2024,hellmann_borehole_2023,phillips_traveltime_1991,zhang_nonlinear_1998,dines_computerized_1979}, and only highlights the similarities and mostly-shared ground between them.

\subsubsection{Basic framework}

Let's assume the environment to be modeled is divided into an $\sz{N_1}{N_2}$ grid of cells (horizontal $\mtimes$ vertical), with a total of $N = N_1\cdot N_2$ cells. A total of $M$ rays are produced, between any number of sources and receivers. We denote $\bvz$ as a $\sz{N}{1}$ positive vector representing the slowness in each cell, and $\bvj{m}(\bvz)$ as a $\sz{N}{1}$ vector containing the length traversed in each cell, by the $m$-th ray. The length-vector $\bvj{m}(\bvz)$ is commonly obtained externally via a fast-marching algorithm or another solver for the Eikonal equation, considering Fermat's principle of least time, and Snell's law $\bvj{m}(\bvz)$. $\bvj{m}(\bvz)$ can also be obtained directly if one considers a slowness profile with small perturbations, where ray-bending effects can be ignored, and the path is assumed linear between source and receiver. With this, the travel-time for the $m$-th ray is $t_m$, being given by
\begin{equation}
	t_m(\bvz) = \tr{\bvj{m}}(\bvz) \bvz
\end{equation}

Now we let $\bvt(\bvz)$ be the $\sz{M}{1}$ vector of travel-times for the current model, for all $M$ considered rays; and $\bvJ(\bvz)$ an $\sz{M}{N}$ matrix, where the $m$-th row is $\tr{j_m}(\bvz)$. With this, our minimization problem becomes minimizing the error $E(\bvz)$, this being translated into
\begin{equation}
	\opt{\bvz} = \arg\min_{\bvz} E(\bvz) = \norm{\bvJ(\bvz) \bvz - \bvt{\obs}}^2
	\label{eq:sec2:def_minimization_bvz}
\end{equation}
where $\bvt{\obs}$ is the observed travel-time for each ray.

Note that this problem is non-linear and recursive: the solution to the ray paths, given by $\bvJ(\bvz)$, depends on the slowness profile $\bvz$. Meanwhile, the profile model depends on the achieved solution for the ray paths, $\bvJ(\bvz)$. Given the non-linearity of the problem, this search is usually done iteratively, following
\begin{subequations}
	\begin{gather}
		\bvz{i+1} = \arg\min_{\bvz{i+1}} E(\bvz{i+1}) = \norm{\bvJ(\bvz{i}) \bvz{i+1} - \bvt{\obs}}^2 
		\label{subeq:sec2:iterative_minimization:eq1}\\
		\nabla E(\bvz{i+1}) = 2\tr{\bvJ}(\bvz{i}) \pts{\bvJ(\bvz{i}) \bvz{i+1} - \bvt{\obs}}
		\label{subeq:sec2:iterative_minimization:eq2}
	\end{gather}
	\label{eqs:sec2:iterative_minimization}
\end{subequations}
in which the gradients are taken with respect to each of the slownesses in $\bvz{i+1}$. Assuming that $\tr{\bvJ}(\bvz{i})\bvJ(\bvz{i})$ is invertible, then the solution to this minimization -- which translates to $\nabla E(\bvz{i+1}) = \bv0$ -- is
\begin{subequations}
	\begin{gather}
		\bvz{i+1} = \bvG[t](\bvz{i}) \bvt{\obs} \\
		\bvG[t](\bvz{i}) = \inv*{\tr{\bvJ}(\bvz{i}) \bvJ(\bvz{i})} \tr{\bvJ}(\bvz{i})
	\end{gather}
	\label{eqs:sec2:inversion_zi+1}
\end{subequations}

Since $\bvJ(\bvz{i})$ is updated repeatedly, so has to be $\bvG[t](\bvz{i})$, and thus the solution is iterative.

\subsubsection{Rank-deficient minimization}
\label{subsubsec:rank-definicient-minimization}

For \cref{eqs:sec2:inversion_zi+1}, it is crucial that $\tr{\bvJ}(\bvz{i})\bvJ(\bvz{i})$ is invertible. Given that $\bvJ$ is $\sz{M}{N}$, then $\tr{\bvJ}(\bvz{i})\bvJ(\bvz{i})$ is $\sz{N}{N}$. It is easy to see that $\rank{\bvJ(\bvz{i})} = N$ and $N \leq M$ are necessary and sufficient for the inversion to be possible. Intuitively, this translates to there being more measurements $M$ than unknowns $N$, that at least $N$ of the measurements are meaningful, and that any cell is represented by at least one ray.

In situations where $\tr{\bvJ}(\bvz{i})\bvJ(\bvz{i})$ isn't full-rank, other procedures should be used for finding $\bvG[t](\bvz{i})$. The most popular ones (\cite{ali_opensource_2019,hormati_robust_2010,zhang_nonlinear_1998,tang_travel_2024}) are techniques derived from the gradient descent method \cite{ruder_overview_2017,haji_comparison_2021}, among which the conjugate gradient \cite{yahaya_descent_2023,hager_survey_2006} stands out. {\color{red} Another option, which hasn't been studied in the context of TTT, is through singular value decomposition \cite{barata_moore_2012,golub_calculating_1965} (SVD)-based inversion \cite{brand_fast_2006,chen_treatment_2006}. Other solvers \cite{carson_new_2017,douglas_solving_2016,ku_novel_2013,beik_iterative_2018} are also available and usable in this application, venues which are yet to be explored.}

With this brief discussion, moving forward we assume that the kernel $\tr{\bvJ(\bvz{i})}\bvJ(\bvz{i})$ -- and its modifications via regularization, blurring, or others -- is invertible; situations where this doesn't apply can be treated with the mentioned techniques.

\subsubsection{Addition of regularization}

Depending on the chosen grid's fineness, it is possible that not every cell has a ray passing through it. This causes both an ill-defined problem, given that this cell's slowness can't be determined by minimizing the error $E(\bvz)$; and a non-smooth solution, which isn't expected for the reconstruction of a physically coherent slowness/velocity profile.

A regularization function can both help solve the ill-definition of the problem, improving the chances of it having a unique solution; and impose some condition on the solution vector, be it smoothness, noise/mismatch reduction, or other metrics. Choosing an adequate regularization can also help generalize the found solution to a continuum, separating the minimization process from the chosen spatial model discretization \cite{zhang_nonlinear_1998,delprat-jannaud_illposed_1993}.

From \cref{eq:sec2:def_minimization_bvz}, we can add a regularization term in the form of $P_{\alpha}(\bvz{i+1})$, such that the iterative form of $\bvz{i+1}$ from \cref{eqs:sec2:iterative_minimization} becomes
\begin{subequations}
	\begin{gather}
		\bvz{i+1} = \arg\min_{\bvz{i+1}} E(\bvz{i+1}) = \norm{\bvJ(\bvz{i}) \bvz{i+1} - \bvt{\obs}}^2 + P_\alpha(\bvz{i+1})\\
		\nabla E(\bvz{i+1}) = 2\tr{\bvJ}(\bvz{i}) \pts{\bvJ(\bvz{i}) \bvz{i+1} - \bvt{\obs}} + \nabla P_{\alpha}(\bvz{i+1})
	\end{gather}
\end{subequations}
with $\alpha$ being a controlling parameter. Assuming that the regularization function can be given by $P_{\alpha}(\bvz{i+1}) = \alpha^2\norm{\bvD\bvz{i+1}}^2$ where $\bvD$ is a square matrix, then the solution is
\begin{equation}
	\bvG[t](\bvz{i}) = \inv*{\tr{\bvJ(\bvz{i})} \bvJ(\bvz{i}) + \alpha^2 \tr{\bvD}\bvD} \tr{\bvJ}(\bvz{i})
\end{equation}
for this inversion step to be possible, it is sufficient that $\bvD$ is a full-rank matrix\footnote{By definition, $\tr{\bvJ}(\bvz{i}) \bvJ(\bvz{i})$ is positive-semidefinite, $\tr{\bvD}\bvD$ is positive-definite if it is full-rank square matrix, and the sum of a positive-semidefinite and a positive-definite is always positive-definite, this being a sufficient condition for invertibility.}.

The regularization matrix $\bvD$ can be chosen to improve a range of metrics. Choosing $\bvD = \Id$ \cite{phillips_traveltime_1991} results in the minimization also considering the root mean-square (RMS) of the slowness vector; this is equivalent to considering the presence in white noise in the measurements, and trading model accurateness for noise rejection. Choosing $\bvD z \approx \nabla^2 s(x,y)$, (discretely) approximating the Laplacian of the (continuous) slowness profile \cite{zhang_nonlinear_1998,ali_opensource_2019}, resulting in a minimization that penalizes the Laplacian of the achieved slowness profile; that is, a solution that also considers the curvature/jumps in the SSP, searching for a smoother profile.

\cite{zhang_nonlinear_1998} also briefly comments on using third-degree derivatives instead of second-degree ones (the latter being the Laplacian regularization), which could also result in a good smoothing on the velocity profile. \cite{hormati_robust_2010} proposes a regularization function based on the orthonormal representation of the slowness distribution, and either the $L_0$ or $L_1$ norms.

{\color{red} Another possibility that is still open for exploration is the use of multiple regularization functions. For example, a leveraged usage of both previously presented ones (white-noise regularization, and Laplacian penalization) can lead to a good smoothness on the slowness profile, while maintaining a guaranteed-inversible kernel. That is, we say that
\begin{equation}
	E(\bvz{i+1}) = \norm{\bvJ(\bvz{i}) \bvz{i+1} - \bvt{\obs}}^2 + \alpha^2\norm{\bvD\bvz{i+1}}^2 + \beta^2 \norm{\bvz{i+1}}^2
\end{equation}
for which
\begin{equation}
	\bvG[t](\bvz{i}) = \inv*{\tr{\bvJ(\bvz{i})} \bvJ(\bvz{i}) + \alpha^2 \tr{\bvD}\bvD + \beta^2 \Id} \tr{\bvJ}(\bvz{i})
\end{equation}
The inverting term is guaranteed to be positive-definite, and therefore the safely inversion can be taken.}
\subsubsection{Convolutional averaging/blurring smoothing}

Another interesting tool that is seen in the literature \cite{phillips_traveltime_1991,meyerholtz_convolutional_1989} is to employ a blur on the gradient. That is, we define that
\begin{equation}
	E(\bvz{i+1}) = \norm{\bvJ[d](\bvz{i}) \bvz[d]{i+1} - \bvt{\obs}}^2 + P_\alpha(\bvz[d]{i+1})
\end{equation}
where $\bvJ[d](\bvz{i}) = \bvJ(\bvz{i})\bvB{i}$ and $\bvz[d]{i+1} = \inv{\bvB{i}} \bvz{i+1}$, and $\bvB{i}$ is the $i$-th iteration's blurring matrix (this being based on a Gaussian blur, or any other averaging kernel, that decreases in radius at each iteration). In this scenario, the solution $\bvz{i+1}$ is
\begin{equation}
	\bvG[t](\bvz{i}) = \bvB{i} \inv*{\tr{\bvB{i}} \tr{\bvJ} \bvJ \bvB{i} + \alpha^2 \tr{\bvD}\bvD} \tr{\bvB{i}} \tr{\bvJ}
\end{equation}

To achieve this, the regularization function has to be adapted, such that $P_{\alpha}(\bvz{i+1}) = \alpha^2 \norm{\bvD \inv{\bvB{i}} \bvz{i+1}}^2$; this means the regularization is applied on the ``reverse blurred'' signal. As stated, the blurring matrix $\bvB{i}$ depends on the iterator $i$. Usually, the blur radius is decreased for increasing $i$, aiming at gradually finding a more and more refined and precise (less blurry) solution.

This convolutional averaging (also called ray broadening \cite{meyerholtz_convolutional_1989,phillips_traveltime_1991}) can help find solutions by scattering the rays and expanding the paths taken to neighboring cells, thus improving in the problem of some cells not being traversed by any rays.

In \cite{ali_opensource_2019} a similar blurring framework is proposed, but this blur is applied directly to the gradient, such that
\begin{equation}
	\nabla E(\bvz{i+1}) = 2 \bvB{k} \tr{\bvJ}(\bvz{i})\pts{\bvJ(\bvz{i}) \bvz{i+1} - \bvt{\obs}} + 2 \alpha^2 \tr{\bvD}\bvD \bvz{i+1}
\end{equation}
Since in the cited paper the solution $\bvz{i+1}$ is found via the conjugate gradient (iterative) method, and not by inversion, the blurring matrix is variable within the conjugate gradient iterative search for the inverse (with iterator $k$), rathen than the Eikonal solving iterative process (iterator $i$). {\color{red} Assuming invertibility, the solution can be directly achieved through
\begin{equation}
	\bvG[t](\bvz{i}) = \inv*{\bvB{i} \tr{\bvJ}(\bvz{i}) \bvJ(\bvz{i}) + \tr{\bvD}\bvD}\bvB{i} \bvJ[T](\bvz{i})
\end{equation}
situation in which the blurring can be lessened on each iteration of $i$, or even kept constant.}

Note that both smoothing frameworks presented require either the presence of a regularization term (otherwise the inverse term can be expanded, and the $\bvB{i}$ terms will cancel); or that the system is sufficiently ill-posed such that a non-inverse minimization is necessary, case in which the smoothing effect doesn't cancel with its inverse.

{\color{red}
\subsubsection{SVD pseudo-inversion}

As briefly commented in \cref{subsubsec:rank-definicient-minimization}, one possibility for the inversion in situations where $\bvJ[T](\bvz{i}) \bvJ(\bvz{i})$ is rank-deficient, is the use of an SVD-based inversion. We write $\bvJ(\bvz{i})$ through its singular value decomposition, as
\begin{equation}
	\bvJ(\bvz{i}) = \bvU \bvSi \bvV[T]
\end{equation}
in which $\bvU \in \re^{\sz{M}{M}}$ and $\bvV \in \re^{\sz{N}{N}}$ are the left- and right-singular vectors of $\bvJ(\bvz{i})$, forming and orthonormal bases; and $\bvSi \in \re^{\sz{M}{N}}$ is a diagonal matrix, given by
\begin{equation}
	\bvSi = \begin{bmatrix}
		\lambda_1 & 0 & \cdots & 0 \\
		0 & \lambda_2 & \cdots & 0 \\
		\vdots & \vdots & \ddots & \vdots \\
		0 & 0 & \cdots & \lambda_N \\
		0 & 0 & \cdots & 0 \\
		\vdots & \vdots & \ddots & \vdots \\
		0 & 0 & \cdots & 0
	\end{bmatrix}
\end{equation}
and $\lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_N$ are the singular values of $\bvJ(\bvz{i})$. Assuming that $\rank{\bvJ(\bvz{i})} = R \leq N$, then the last $N-R$ singular values of $\bvJ(\bvz{i})$ -- and thus the last $N-R$ diagonal entries of $\bvSi$ -- will be zero. In this scenario, $\pinv*{\bvJ[T](\bvz{i}) \bvJ(\bvz{i})}$ -- the pseudo-inverse of ${\bvJ[T](\bvz{i}) \bvJ(\bvz{i})}$ -- can be obtained as
\begin{equation}
	\begin{split}
		\pinv*{\bvJ[T](\bvz{i}) \bvJ(\bvz{i})}
		& = \pinv*{\bvV \bvSi[T] \bvU[T] \bvU \bvSi \bvV[T]} \\
		& = \pinv*{\bvV \bvSi[T] \bvSi \bvV[T]} \\
		& = \bvV \pinv*{\bvSi[T] \bvSi} \bvV[T] \\
		& = \bvV \pinv{\bvSi} \pinv{{\bvSi[T]}} \bvV[T]
	\end{split}
\end{equation}
where the orthonormality of $\bvV$ and the distributive property of the pseudo-inverse (considering the orthonormality of $\bvV$ and of the columns of $\bvSi$) were used; and $\pinv{\bvSi} \in\re^{\sz{N}{M}}$ is a diagonal matrix, with its entries being the reciprocal of the singular values of $\bvJ(\bvz{i})$.

With this, the solution to the minimization problem is
\begin{equation}
	\begin{split}
		\bvG[t](\bvz{i})
		& = \pinv*{\bvJ[T](\bvz{i}) \bvJ(\bvz{i})} \bvJ[T](\bvz{i}) \\
		& = \bvV \pinv{\bvSi} \pinv{{\bvSi[T]}} \bvV[T] \bvV \bvSi[T] \bvU[T] \\
		& = \bvV \pinv{\bvSi} \bvU[T]
	\end{split}
\end{equation}

With this solution, the gradient for $\bvz{i} = \bvV \pinv{\bvSi} \bvU[T] \bvt$ is
\begin{equation}
	\begin{split}
		\nabla E(\bvz{i})
		& = 2\pts{\bvV \bvSi[T] \bvU[T] \bvU \bvSi \bvV[T] \bvV \pinv{\bvSi} \bvU[T] \bvt - \bvV \bvSi[T] \bvU[T] \bvt} \\
		& = 2\pts{\bvV \bvSi[T] \bvSi \pinv{\bvSi} \bvU[T] \bvt - \bvV \bvSi[T] \bvU[T] \bvt} \\
		& = 2\pts{\bvV \bvSi[T] \bts{\Id{M;R} - \Id{M}} \bvU[T] \bvt}
	\end{split}
\end{equation}
where $\Id{M;R} \in \re^{\sz{M}{M}}$ is a diagonal matrix with the first $R$ entries being $1$, and the last $M-R$ ones being $0$; and $\Id{M} \in \re^{\sz{M}{M}}$ is the identity matrix. From this, $\Id{M;R} - \Id{M}$ has its first $R$ diagonal entries $0$, and the last ones $-1$. Since all rows of $\bvSi[T]$ are in the null space of $\Id{M;R} - \Id{M}$ (only the first $R$ entries of any row of $\bvSi[T]$ are non-zero, exactly where $\Id{M;R} - \Id{M}$ is zero), then $\bvSi[T]\bts{\Id{M;R} - \Id{M}} = \bv0$. Therefore,
\begin{equation}
	\begin{split}
		\nabla E(\bvV \pinv{\bvSi} \bvU[T] \bvt)
		& = 2\pts{\bvV \,\bv0\, \bvU[T] \bvt} \\
		& = \bv0
	\end{split}
\end{equation}

With this, we show that $\bvz{i+1} = \bvV \pinv{\bvSi} \bvU[T] \bvt$ is a solution to the minimization process.

Furthermore, if we write
\begin{equation}
	\bvV = \begin{bmatrix}
		\bvV{1} & \bvV{2}
	\end{bmatrix}
\end{equation}
where $\bvV{1}\in \re^{\sz{N}{R}}$ are the first $R$ columns of $\bvV$, and $\bvV{2} \in \re^{\sz{N}{(N-R)}}$ the remaining $N-R$ columns of $\bvV$, relative to the zero rows of $\bvSi$; that is, the columns of $\bvV{2}$ form an orthonormal basis for the null space of $\bvSi$ (and thus, of $\bvJ(\bvz{i})$).

For any $\bvmu \in \re^{\sz{(N-R)}{1}}$, any vector of the form $\bvz[t] = \bvV{2} \bvmu$ will be within the null space of $\bvSi$, and thus $\bvJ \bvz[t] = \bvU \bvSi \bvV[T] \bvV{2} \bvmu$ will be identically $0$. Therefore,
\begin{equation}
	\nabla E(\bvV \pinv{\bvSi} \bvU[T] \bvt + \bvV{2} \bvmu) = \bv0~\forall~\bvmu \in \re^{\sz{(N-R)}{1}}
\end{equation}

This information could be leveraged to find a solution that is optimal in another metric. For example, we can define a new cost function $F(\bvmu)$, which is given by
\begin{equation}
	F(\bvmu) = \norm{\bvD \pts{\bvV \pinv{\bvSi} \bvU{T} \bvt + \bvV{2}\bvmu}}^2
\end{equation}
where $\bvD$ is the previously determined discrete approximation for the Laplacian. Since $\rank{\bvV{2}} = N-R$, given that all of its columns are orthonormal, and assuming that $\bvD$ is a full-rank matrix, a solution to this problem can be directly found, this being given by
\begin{equation}
	\bvmu = -\inv*{\bvV[T]{2} \bvD[T] \bvD\bvV{2}} \bvV[T]{2} \bvD[T] \bvD \bvV \pinv{\bvSi} \bvU[T] \bvt
\end{equation}

Defining $\bvP \bvOm \bvQ[T]$ as the singular value decomposition of $\bvD \bvV{2}$, then
\begin{equation}
	\bvmu = -\bvQ \inv{\bvOm} \bvP[T] \bvD \bvV \pinv{\bvSi} \bvU[T] \bvt
\end{equation}

This solution both has zero error between the measured and estimated travel-times, and has minimum Laplacian within the space of zero-error solutions. The main insight for this is to exploit the null-space of $\bvJ$, and minimize the main objective within the row space, and minimize the second objective within the null space. Note that this procedure only works if $\rank{\bvJ(\bvz{i})} = R < N$. If $R = N$, there is no null space to be exploited, and therefore the second minimization can't be done.
}
\newpage
