\section{Fundamentals of travel-time tomography}

As previously explained, Travel-Time Tomography (TTT) uses the delay -- travel time -- between the wave's generation at the source and its arrival at the receptors, to estimate the path traveled, and therefore the wave's speed along the route taken.

\input{figures/1.1-WavepathHeterogeneous.tex}

Consider a wave that follows the path from \cref{fig:sec2:wavepath_heterogeneous-variable_environment}, this being represented by the curve $\bvr[v]$. Assuming that the speed of sound $c(\bvp)$ is a function of space ($\bvp = [x,y,z]$), then the travel time along the curve $\bvr[v]$ is
\begin{equation}
	t_{\bvr[v]} = \int_{\bvr[v]} \frac{1}{c(\bvp)} \dd\bvp.
\end{equation}

The objective is to, given a set of $N$ measures $t_{\bvr[v]{i}}$, these representing different wave paths, estimate the sound speed field $c(\bvp)$.

Different approaches can be taken to solve this problem, including to discretize space, approximate the SSF as polynomial through Spline or least-squares fitting, among others. This technique can also be employed in 1D, 2D or 3D situations, and acquiring spatial diversity through either (or both) multi-source or multi-receiver means.

Most models estimate the slowness (inverse of speed) field. Note that SSF can both mean \quote{sound speed field} and \quote{sound slowness field}. Since both quantities are the reciprocal of one another, the SSF acronym will be indiscreetly used for both.

\subsection{Papers and articles' review}

As stated previously, this technique is versatile due to its simplicity, being employable in a range of applications from acoustics seismology to computarized living-tissue tomography and electromagnetic evaluation of water properties.

\subsubsection{Aki, K - 1976, 1977}

The works by Aki in \cite{aki_determination_1976,aki_determination_1977} are sometimes regarded as the pioneering papers in travel-time tomography. Their work is in the field of seismology and geological layer classification through the SSF estimation.

\textbf{Objective} - The purpose of this paper is to use travel-time information at different geologic layers to estimate the SSF in them, as well as guiding earthquake origin/epicenter estimation.

\textbf{Data and assumptions} - The data used is the travel-time from sub-surface waves generated by earthquakes. For such, the media (geologic layers) are discretized among them, as well as each layer being separated into cells, with varying velocities within each cell. The environment is assumed at least 2D, as the media is separated into layers (vertical quantization) and then into cells (horizontal discretization of the said layers).

\textbf{Techniques used} - Each layer is assumed to have an average velocity, and each cell in that layer deviates from this average. Least-squares technique is used: the model is adjusted to match the modeled and measured first arrival times.

\subsubsection{Ali - 2019}

This paper \cite{ali_opensource_2019} aims to estimate the SSF on different tissues via ultrasound computer tomography. They use a joint Eikonal equation and ray tracing solutions, coupled with different regularization techniques for smoothness constraints.

\textbf{Objective} - Use travel-time information to reconstruct the SSF, allowing for a better imaging process for biological tissue.

\textbf{Data and assumptions} - Travel-time information from ultrasound CTs, with a ring of receivers around the object.

\textbf{Technique used} - The tissue is considered to be split into 2D cells. An iterative minimization process is employed, searching concomitantly for the SSF of the tissue, as well as the path taken by the emitted signals through the tissue. The Fast Marching Method (FMM) \cite{chopp_improvements_2001,sethian_3d_1999,sethian_fast_1999} is used for the raytracing portion, and an iterative least-squares minimization to find most appropriate speed on each cell in the tissue.

\textbf{Similar paper} - The paper by Hormati \cite{hormati_robust_2010} uses a similar approach to that in \cite{ali_opensource_2019}. Their objective is to determine the SSF in an ultrasound tomography by: discretizing the tissue; applying a (discrete version of) the travel-time integral to the system; and estimating the system parameters, as well as the ray-lengths within each discrete cell, to iteratively achieve the tissue's SSF. They apply a few different techniques for the iterative minimization process, but this contrast is outside the TTT-SSF scope.

\textbf{Similar paper} - The paper by Jeong \cite{jeong_investigating_2024} uses a similar approach to that in \cite{hormati_robust_2010}, comparing TTT to reflection tomography, and employing them in convolutional neural network (CNN) deep learning training.

\textbf{Similar paper} - In this paper by Hellmann \cite{hellmann_borehole_2023}, they use travel-time tomography for estimating the SSF in glacier boreholes, along with a model correction step for the positions of the sources/receivers along the boreholes.

\subsubsection{Philips - 1991}

This paper \cite{phillips_traveltime_1991} studies different methods for discrete travel-time tomography, comparing their results in metrics such as roughness, misfit, and spread. These methods trade overall performance against that of some metric.

\textbf{Objective} - Compare different inverse paradigms in achieving the parameters for the SSF model.

\textbf{Data and assumptions} - Synthetic discretized data. There are 100 sensors/receivers in total, 25 per side of the square-grid of sensors. Each device sends array to all other devices. Ray-bending is disregarded.

\textbf{Technique used} - Similar to the previously studied papers, the 2D environment is split into cells, and inversion of the slowness model is used to achieve the results. Different conditions on the inversion are considered, such as smoothing, regularization, dampening, and iterative reweighting. These different techniques are compared for the proposed metrics, exploring their advantages for different applications.

\subsubsection{Zhang - 2017}

This paper \cite{zhang_nonlinear_1998} proposes an adaptation of the travel-time tomography approach to include reflections as well. Their modeling minimize a combination of average slowness (travel-time $/$ path length) and apparent slowness (derivative of travel-time w.r.t. surface (horizontal) distance), along with a regularization term.

\textbf{Objective} - Achieve a TTT model that incorporates reflections alongside direct-path waves, in estimating the parameters of the system.

\textbf{Data and assumptions} - It assumes that reflected waves can arrive in the sensors at similar times as the direct-path waves, enabling their coupled utilization in estimating the system.

\textbf{Technique used} - A conjugate gradient minimization approach is employed, as is usual in TTT inversion problems. Also, similarly to previously explored papers, a cell-based approach is taken.

\subsubsection{Dines - 1979}

This paper \cite{dines_computerized_1979} also works in TTT, now estimating both the velocity and attenuation fields from the measured data. This paper works with electromagnetic signals, however the mathematics and theories are the same.

\textbf{Data and assumptions} - They assume an exponential decay on the signals' amplitudes over time, as well as straight waves, dismissing refraction effects. The same media discretization is done, separating it into 2D cells along the depth and distance between the columns of sources/receivers.

\textbf{Technique used} - An iterative minimization is done over all sources and sensors (and their pair-wise paths), searching for either the velocity or attenuation fields. For solving the ill-posed problem, they propose an approach that considers one ray at a time, updating all velocities in the cells that are traversed by the current ray, until sufficient convergence is achieved.

\subsubsection{Tang - 2024}

This paper \cite{tang_travel_2024} proposes the use of the Fast Sweeping Method (FSM) instead of FMM for solving the Eikonal equations during the TTT inversion process. They present the FSM method and derive its implementation for TTT, as well as an extensive comparison between the two solvers for the Eikonal problem. Modeling of reflections is also discussed in the paper.

%\textbf{Objective} - Implement a FSM-based TTT technique for SSF estimation.

\textbf{Data and assumptions} - Synthetic environment and signals are considered, for a wide gamma of situations.

\textbf{Technique used} - The paper's focus is on exploring the use of FSM instead of FMM, and not so much in the TTT process, or other details surrounding it.

\subsubsection{Common themes between articles}
\NewDocumentCommand{\citep}{m}{%
	\citeauthor*{#1}, \citeyear{#1}%
}
\def\X{\ensuremath{\mathsf{X}}}

The table below compiles a comprehensive list of the themes that are commonly approached among the papers presented previously, as well as a short commentary on the different features or considerations, when relevant.
\begin{table}[H]\small\centering
	\begin{tabular}{l c c c c c l}
		Paper 								& Discrete 	& Ray bending 	& CG 	& $P_{\alpha}$ 	& $\bvB{i}$ & Extra 	\\
		\citep{aki_determination_1976} 		& \X 		& -- 			& --	& \X			& --		& --		\\
		\citep{aki_determination_1977} 		& \X 		& -- 			& --	& \X			& --		& Perturbation analysis 		\\
		\citep{ali_opensource_2019}			& \X		& \X			& \X	& \X			& \X		& Presented a statistical linearization \\
		\citep{hormati_robust_2010}			& \X 		& \X			& \X	& \X			& --		& --		\\
		\citep{jeong_investigating_2024}	& \X		& \X			& --	& \X			& --		& --		\\
		\citep{hellmann_borehole_2023}		& \X		& \X			& --	& \X			& --		& Position estimation	\\
		\citep{phillips_traveltime_1991}	& \X		& --			& --	& \X			& \X		& Compares many techniques	\\
		\citep{zhang_nonlinear_1998}		& \X		& \X			& \X	& \X			& --		& Considers reflections	\\
		\citep{dines_computerized_1979}		& \X		& --			& --	& --			& --		& Considers attenuation; Works ray-by-ray \\
		\citep{tang_travel_2024}			& \X		& \X			& \X	& --			& --		& Uses FSM instead of FMM
	\end{tabular}
\end{table}

\subsection{Standard discrete travel-time tomography approach}
\label{subsec:sec2:standard_ttt_approach}

All of the cited papers took an approach similar to the one established below. Differences may arise from ignoring ray bending, type of regularization in the minimization process, amount and positioning of sources and receivers, final purpose for the TTT, or some other particularization or assumption.

In these regards, the most complete papers are the ones by Ali \cite{ali_opensource_2019}, Phillips \cite{phillips_traveltime_1991}, and Hormati \cite{hormati_robust_2010}, where they consider regularization, ray bending, and a large amount of sources and receivers. We now will lay the common ground among all the presented papers, the median modeling considerations and formulations, as well as some important distinctions and particularities that are shared between some of them. Note that the framework presented here isn't identical to the ones in the cited papers \cite{aki_determination_1977,aki_determination_1976,ali_opensource_2019,hormati_robust_2010,jeong_investigating_2024,hellmann_borehole_2023,phillips_traveltime_1991,zhang_nonlinear_1998,dines_computerized_1979}, and only highlights the similarities and mostly-shared ground between them.

\subsubsection{Basic framework}

Let's assume the environment to be modeled is divided into an $\sz{N_1}{N_2}$ grid of cells (horizontal $\mtimes$ vertical), with a total of $N = N_1\cdot N_2$ cells. A total of $M$ rays are produced, between any number of sources and receivers. We denote $\bvz \in \re^{\sz{N}{1}}$ positive vector representing the slowness in each cell, and $\bvj{m}(\bvz) \in \re^{\sz{N}{1}}$ vector containing the length traversed in each cell, by the $m$-th ray. The length-vector $\bvj{m}(\bvz)$ is commonly obtained externally via a fast-marching algorithm or another solver for the Eikonal equation, considering Fermat's principle of least time, and Snell's law $\bvj{m}(\bvz)$. $\bvj{m}(\bvz)$ can also be obtained directly if one considers an SSF with small perturbations, where ray-bending effects can be ignored, and the path is assumed linear between source and receiver. With this, the travel-time for the $m$-th ray is $t_m$, being given by
\begin{equation}
	t_m(\bvz) = \tr{\bvj{m}}(\bvz) \bvz
\end{equation}

Now we let $\bvt(\bvz) \in \re^{\sz{M}{1}}$ vector of travel-times for the current model, for all $M$ considered rays; and $\bvJ(\bvz) \in \re^{\sz{M}{N}}$ matrix, where the $m$-th row is $\tr{j_m}(\bvz)$. With this, our minimization problem becomes minimizing the error $E(\bvz)$, this being translated into
\begin{equation}
	\opt{\bvz} = \arg\min_{\bvz} E(\bvz) = \norm{\bvJ(\bvz) \bvz - \bvt{\obs}}^2
	\label{eq:sec2:def_minimization_bvz}
\end{equation}
where $\bvt{\obs}$ is the observed travel-time for each ray.

Note that this problem is non-linear and recursive: the solution to the ray paths, given by $\bvJ(\bvz)$, depends on the slowness field $\bvz$. Meanwhile, the field model depends on the achieved solution for the ray paths, $\bvJ(\bvz)$. Given the non-linearity of the problem, this search is usually done iteratively, following
\begin{subequations}
	\begin{gather}
		\bvz{i+1} = \arg\min_{\bvz{i+1}} E(\bvz{i+1}) = \norm{\bvJ(\bvz{i}) \bvz{i+1} - \bvt{\obs}}^2 
		\label{subeq:sec2:iterative_minimization:eq1}\\
		\nabla E(\bvz{i+1}) = 2\tr{\bvJ}(\bvz{i}) \pts{\bvJ(\bvz{i}) \bvz{i+1} - \bvt{\obs}}
		\label{subeq:sec2:iterative_minimization:eq2}
	\end{gather}
	\label{eqs:sec2:iterative_minimization}
\end{subequations}
in which the gradients are taken with respect to each of the slownesses in $\bvz{i+1}$. Assuming that $\tr{\bvJ}(\bvz{i})\bvJ(\bvz{i})$ is invertible, then the solution to this minimization -- which translates to $\nabla E(\bvz{i+1}) = \bv0$ -- is
\begin{subequations}
	\begin{gather}
		\bvz{i+1} = \bvG(\bvz{i}) \bvt{\obs} \\
		\bvG(\bvz{i}) = \inv*{\tr{\bvJ}(\bvz{i}) \bvJ(\bvz{i})} \tr{\bvJ}(\bvz{i})
	\end{gather}
	\label{eqs:sec2:inversion_zi+1}
\end{subequations}

Since $\bvJ(\bvz{i})$ is updated repeatedly, so has to be $\bvG(\bvz{i})$, and thus the solution is iterative. Although $\bvJ(\bvz{i})$ and $\bvG(\bvz{i})$ directly depend on the acquired discrete slowness fields (via the FMM or FSM solvers), the dependence on the iterator $i$ is much more relevant. For this, and for economy of notation, we let $\bvJ{i} \equiv \bvJ(\bvz{i})$, and $\bvG{i} = \bvG(\bvz{i})$.

\subsubsection{Rank-deficient minimization}
\label{subsubsec:rank-definicient-minimization}

For \cref{eqs:sec2:inversion_zi+1}, it is crucial that $\bvJ[T]{i}\bvJ{i}$ is invertible. Given that $\bvJ{i}  \in \re^{\sz{M}{N}}$, then $\bvJ[T]{i}\bvJ{i}  \in \re^{\sz{N}{N}}$. It is easy to see that $\rank{\bvJ{i}} = N$ and $N \leq M$ are necessary and sufficient for the inversion to be possible. Intuitively, this translates to there being more measurements $M$ than unknowns $N$, that at least $N$ of the measurements are meaningful (linearly independent), and that any cell is represented by at least one ray.

In situations where $\bvJ[T]{i}\bvJ{i}$ isn't full-rank, other procedures should be used for finding $\bvG{i}$. The most popular ones (used by \cite{ali_opensource_2019,hormati_robust_2010,zhang_nonlinear_1998,tang_travel_2024}) are techniques derived from the gradient descent method \cite{ruder_overview_2017,haji_comparison_2021}, among which the conjugate gradient \cite{yahaya_descent_2023,hager_survey_2006} stands out. {\proposed Another option, which hasn't been studied in the context of TTT, is through singular value decomposition (SVD) \cite{barata_moore_2012,golub_calculating_1965} -based inversion \cite{brand_fast_2006,chen_treatment_2006}. Other solvers \cite{carson_new_2017,douglas_solving_2016,ku_novel_2013,beik_iterative_2018} are also available and usable in this application, venues which are yet to be explored.}

With this brief discussion, moving forward we assume that the kernel $\tr{\bvJ{i}}\bvJ{i}$ -- and its modifications via regularization, blurring, or others -- is invertible; situations where this doesn't apply can be treated with the mentioned techniques.

\subsubsection{Addition of regularization}

Depending on the chosen grid's fineness, it is possible that not every cell has a ray passing through it. This causes both an ill-defined problem, given that this cell's slowness can't be determined by minimizing the error $E(\bvz)$; and a non-smooth solution, which isn't expected for the reconstruction of a physically coherent slowness/velocity field.

A regularization function can both help solve the ill-definition of the problem, improving the chances of it having a unique solution; and impose some condition on the solution vector, be it smoothness, noise/mismatch reduction, or other metrics. Choosing an adequate regularization can also help generalize the found solution to a continuum, separating the minimization process from the chosen spatial model discretization \cite{zhang_nonlinear_1998,delprat-jannaud_illposed_1993}.

From \cref{eq:sec2:def_minimization_bvz}, we can add a regularization term in the form of $P_{\alpha}(\bvz{i+1})$, such that the iterative form of $\bvz{i+1}$ from \cref{eqs:sec2:iterative_minimization} becomes
\begin{subequations}
	\begin{gather}
		\bvz{i+1} = \arg\min_{\bvz{i+1}} E(\bvz{i+1}) = \norm{\bvJ{i} \bvz{i+1} - \bvt{\obs}}^2 + P_\alpha(\bvz{i+1})\\
		\nabla E(\bvz{i+1}) = 2\bvJ[T]{i} \pts{\bvJ{i} \bvz{i+1} - \bvt{\obs}} + \nabla P_{\alpha}(\bvz{i+1})
	\end{gather}
	\label{eqs:sec2:iterative_minimization_Ezi}
\end{subequations}
with $\alpha$ being a controlling parameter. Assuming that the regularization function can be given by $P_{\alpha}(\bvz{i+1}) = \alpha^2\norm{\bvD\bvz{i+1}}^2$ where $\bvD$ is a square matrix, then the solution is
\begin{equation}
	\bvG{i} = \inv*{\tr{\bvJ{i}} \bvJ{i} + \alpha^2 \tr{\bvD}\bvD} \bvJ[T]{i}
	\label{eq:sec2:solution_regularization-problem_simple}
\end{equation}
for this inversion step to be possible, it is sufficient that $\bvD$ is a full-rank matrix\footnote{\label{footnote:positive-definite_inversion}By definition, $\bvJ[T]{i} \bvJ{i}$ is positive-semidefinite, $\tr{\bvD}\bvD$ is positive-definite if it is full-rank square matrix, and the sum of a positive-semidefinite and a positive-definite is always positive-definite, this being a sufficient condition for invertibility.}.

The regularization matrix $\bvD$ can be chosen to improve a range of metrics. Choosing $\bvD = \Id$ \cite{phillips_traveltime_1991} results in the minimization also considering the root mean-square (RMS) of the slowness vector; this is equivalent to considering the presence in white noise in the measurements, and trading model accurateness for noise rejection. Choosing $\bvD z \approx \nabla^2 s(x,y)$, (discretely) approximating the Laplacian of the (continuous) slowness field \cite{zhang_nonlinear_1998,ali_opensource_2019}, resulting in a minimization that penalizes the Laplacian of the achieved slowness field; that is, a solution that also considers the curvature/jumps in the SSF, searching for a smoother field.

\cite{zhang_nonlinear_1998} also briefly comments on using third-degree derivatives instead of second-degree ones (the latter being the Laplacian regularization), which could also result in a good smoothing on the velocity field. \cite{hormati_robust_2010} proposes a regularization function based on the orthonormal representation of the slowness distribution, and either the $L_0$ or $L_1$ norms.

{\proposed Another possibility that is still open for exploration is the use of multiple regularization functions. For example, a leveraged usage of both previously presented ones (white-noise regularization, and Laplacian penalization) can lead to a good smoothness on the SSF, while maintaining a guaranteed-invertible kernel. That is, we say that
	\begin{equation}
		E(\bvz{i+1}) = \norm{\bvJ{i} \bvz{i+1} - \bvt{\obs}}^2 + \alpha^2\norm{\bvD\bvz{i+1}}^2 + \beta^2 \norm{\bvz{i+1}}^2
	\end{equation}
	for which
	\begin{equation}
		\bvG{i} = \inv*{\tr{\bvJ{i}} \bvJ{i} + \alpha^2 \tr{\bvD}\bvD + \beta^2 \Id} \bvJ[T]{i}
	\end{equation}
	
	The brackets term is guaranteed to be positive-definite (\cref{footnote:positive-definite_inversion}), and therefore safely invertible.
}
\subsubsection{Convolutional averaging/blurring smoothing}

Another interesting tool that is seen in the literature \cite{phillips_traveltime_1991,meyerholtz_convolutional_1989} is to employ a blur on the gradient. That is, we define that
\begin{equation}
	E(\bvz{i+1}) = \norm{\bvJ[d]{i} \bvz[d]{i+1} - \bvt{\obs}}^2 + P_\alpha(\bvz[d]{i+1})
\end{equation}
where $\bvJ[d]{i} = \bvJ{i}\bvB{i}$ and $\bvz[d]{i+1} = \inv{\bvB{i}} \bvz{i+1}$, and $\bvB{i}$ is the $i$-th iteration's blurring matrix (this being based on a Gaussian blur, or any other averaging kernel, that decreases in radius at each iteration). In this scenario, the solution $\bvz{i+1}$ is
\begin{equation}
	\bvG{i} = \bvB{i} \inv*{\tr{\bvB{i}} \tr{\bvJ{i}} \bvJ{i} \bvB{i} + \alpha^2 \tr{\bvD}\bvD} \tr{\bvB{i}} \bvJ[T]{i}
\end{equation}

To achieve this, the regularization function has to be adapted, such that $P_{\alpha}(\bvz{i+1}) = \alpha^2 \norm{\bvD \inv{\bvB{i}} \bvz{i+1}}^2$; this means the regularization is applied on the ``reverse blurred'' signal. As stated, the blurring matrix $\bvB{i}$ depends on the iterator $i$. Usually, the blur radius is decreased for increasing $i$, aiming at gradually finding a more refined and accurate (less blurry) solution.

This convolutional averaging (also called ray broadening \cite{meyerholtz_convolutional_1989,phillips_traveltime_1991}) can help find solutions by scattering the rays and expanding the paths taken to neighboring cells, thus dealing with the problem of some cells not being traversed by any rays.

In \cite{ali_opensource_2019} a similar blurring framework is proposed, but this blur is applied directly to the gradient, such that
\begin{equation}
	\nabla E(\bvz{i+1}) = 2 \bvB{k} \bvJ[T]{i}\pts{\bvJ{i} \bvz{i+1} - \bvt{\obs}} + 2 \alpha^2 \tr{\bvD}\bvD \bvz{i+1}
\end{equation}
Since in the cited paper the solution $\bvz{i+1}$ is found via the conjugate gradient (iterative) method, and not by inversion, the blurring matrix is variable within the conjugate gradient iterative search for the inverse (with iterator $k$), rathen than the Eikonal solving iterative process (iterator $i$). {\proposed Assuming invertibility, the solution can be directly achieved through
	\begin{equation}
		\bvG{i} = \inv*{\bvB{i} \bvJ[T]{i} \bvJ{i} + \tr{\bvD}\bvD}\bvB{i} \bvJ[T]{i}
	\end{equation}
	situation in which the blurring can be lessened on each iteration of $i$, or even kept constant.}

Note that both smoothing frameworks presented require either the presence of a regularization term (otherwise the inverse term can be expanded, and the $\bvB{i}$ terms will cancel); or that the system is sufficiently ill-posed such that a non-inverse minimization is necessary, case in which the smoothing effect doesn't necessarily cancel with its inverse.